---
title: "Seffane VE"
author: "SEFFANE Asmaa"
date: "2023-07-31"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
INTRODUCTION:

Extreme value theory is concerned with the statistical properties of the extreme events related to a random variable and the understanding and applications of their probability distributions.

In this project, I will study two data:

Fort Collins (Colorado)
Daily precipitation amounts (inches) from a single rain gauge in Fort Collins, Colorado.
The variable of interest is the daily precipitation amount (inches) : Prec.

Dowjones:
This data shows the daily closing prices of the Dow Jones Index over a 5-year period.
Evidently, the level of the process has changed dramatically over the observed period, and issues about extremes of daily behavior are swamped by long-term time variation in the series.

Fort Collins Colorado:

Descriptives statistics:

I start  by requiring all the packages i will need:

```{r}
require(evd)
require(evir)
require(ismev)
require(fExtremes)
require(extRemes)
```

```{r}
data(Fort)
str(Fort$Prec)
length(Fort$Prec)
summary(Fort)
plot(Fort$Prec, col = "blue",pch=16)
boxplot(Fort$Prec)
hist(Fort$Prec, col = "purple")
emplot(Fort$Prec, col = "blue")
```
As it's required, I have to use both approaches (GEV and GPD) and compare them.

The Generalized Extreme Value distribution : GEV

The block maxima approach:

```{r}
bmFort<- blockmaxxer(Fort, blocks = Fort$year, which="Prec")
head(bmFort)
length(as.numeric(rownames(bmFort)))
plot(Fort$Prec,ylab = "Prec",col = "blue",pch = 16)
points(as.numeric(rownames(bmFort)),bmFort$Prec, col="red", pch=16)
```

Maximum Likelihood Estimators:
```{r}
FitFortMle = fevd(bmFort$Prec, type="GEV", method="MLE")
FitFortMle
ci(FitFortMle,type="parameter")

```

```{r}
plot(FitFortMle,type="qq")
plot(FitFortMle,type="qq2")
plot(FitFortMle,type="density")
plot(FitFortMle,type="rl")

```


I will try the Gumbel model:

```{r}
FitFortGumbelMle=fevd(bmFort$Prec,type="Gumbel",method="MLE")
FitFortGumbelMle
ci(FitFortGumbelMle)
```

```{r}
plot(FitFortGumbelMle,type="qq")
plot(FitFortGumbelMle,type="qq2")
plot(FitFortGumbelMle,type="density")
plot(FitFortGumbelMle,type="rl")

```


COMPARAISON:

```{r}
lr.test(FitFortMle,FitFortGumbelMle)
```
Probability Weighted Moment : PWM

```{r}
FitFortMoments=fevd(bmFort$Prec,type="GEV",method="Lmoments")
FitFortMoments
ci(FitFortMoments,type="parameter")
```
```{r}
plot(FitFortMoments)
```

Return Level for Fort GEV MLE and PWM:

```{r}
return.level(FitFortMle, return.period = c(100, 1000))

return.level(FitFortGumbelMle, return.period = c(100, 1000))

return.level(FitFortMoments, return.period = c( 100, 1000))

```

```{r}
summary(Fort$Prec)
```

The Generalized Pareto Distribution:

```{r}
plot(Fort$Prec,type="h",col="blue")
points(as.numeric(rownames(bmFort)),bmFort$Prec, col="red", pch=16)

```


```{r}
plot(Fort$Prec,type="h",col="blue")
abline(h=1, col="red",lwd=2)
```


```{r}
length(Fort$Prec > 2)

```

```{r}
plot(Fort$Prec,type="h",col="blue")
abline(h=2, col="red",lwd=2)
```

```{r}
 mrlplot(Fort$Prec)
```



```{r}
threshrange.plot(Fort$Prec, r = c(1, 2), nint=10)
```
1.2 or 1.5

I choose 1.2
Maximum Likelihood Estimators : MLE
```{r}
FitFortgpdMle =fevd(Fort$Prec,threshold=1.2,type="GP",method="MLE",time.units="days")
FitFortgpdMle
ci(FitFortgpdMle,type="parameter")
```

```{r}
plot(FitFortgpdMle,type="qq")
plot(FitFortgpdMle,type="qq2")
plot(FitFortgpdMle,type="density")
plot(FitFortgpdMle,type="rl")
```

```{r}
FitFortgpdMleExp=fevd(Fort$Prec,threshold=1.2,type="Exponential",method="MLE",time.units="days")
FitFortgpdMleExp
ci(FitFortgpdMleExp,type="parameter")
```


```{r}
plot(FitFortgpdMleExp,type="qq")
plot(FitFortgpdMleExp,type="qq2")
plot(FitFortgpdMleExp,type="density")
plot(FitFortgpdMleExp,type="rl")
```

Probability Weighted Moment PWM:

```{r}
FitFortgpdMoments=fevd(Fort$Prec,threshold=1.2,type="GP",method="Lmoments",time.units="days")
FitFortgpdMoments
ci(FitFortgpdMoments,type="parameter")

```

```{r}
plot(FitFortgpdMoments,type="qq")
plot(FitFortgpdMoments,type="qq2")
plot(FitFortgpdMoments,type="density")
plot(FitFortgpdMoments,type="rl")
```

Return Level:
```{r}
return.level(FitFortgpdMle, return.period = c(20, 100, 1000))

return.level(FitFortgpdMleExp, return.period = c(20, 100, 1000))

return.level(FitFortgpdMoments, return.period = c(20, 100, 1000))

```

COMPARAISON:

COMPARAISON between GEV and GPD:

CONCLUSION:


DATA OF portpirie

```{r}
data(portpirie)
str(portpirie)
length(portpirie)
summary(portpirie)
plot(portpirie$SeaLevel, col = "blue", pch = 16)
boxplot(portpirie$SeaLevel)
hist(portpirie$SeaLevel, col = "purple")
emplot(portpirie$SeaLevel, col = "blue")
```


The Generalized Extreme Value distribution : GEV

The block maxima approach:

```{r}
bmSea<- blockmaxxer(portpirie, blocks = portpirie$Year, which="SeaLevel")
head(bmSea)
as.numeric(rownames(bmSea))
length(as.numeric(rownames(bmSea)))
plot(portpirie$SeaLevel,ylab = "Sealevel",col = "blue",pch = 16)
points(as.numeric(rownames(bmSea)),portpirie$SeaLevel, col="red", pch=16)
```

Maximum Likelihood Estimators:
```{r}
FitSeaMle = fevd(portpirie$SeaLevel, type="GEV", method="MLE")
FitSeaMle
ci(FitSeaMle,type="parameter")
```

```{r}
plot(FitSeaMle,type="qq")
plot(FitSeaMle,type="qq2")
plot(FitSeaMle,type="density")
plot(FitSeaMle,type="rl")
```




I will try the Gumbel model:

```{r}
FitSeaGumbelMle=fevd(portpirie$SeaLevel,type="Gumbel",method="MLE")
FitSeaGumbelMle
ci(FitSeaGumbelMle)
```



```{r}
plot(FitSeaGumbelMle,type="qq")
plot(FitSeaGumbelMle,type="qq2")
plot(FitSeaGumbelMle,type="density")
plot(FitSeaGumbelMle,type="rl")
```

COMPARAISON:

```{r}
lr.test(FitSeaMle,FitSeaGumbelMle)
```
Probability Weighted Moment : PWM

```{r}
FitSeaMoments=fevd(portpirie$SeaLevel,type="GEV",method="Lmoments")
FitSeaMoments
ci(FitSeaMoments,type="parameter")
```

```{r}
plot(FitSeaMoments)
```

Return Level for Fort GEV MLE and PWM:

```{r}
return.level(FitSeaMle, return.period = c(100, 1000))

return.level(FitSeaGumbelMle, return.period = c(100, 1000))

return.level(FitSeaMoments, return.period = c( 100, 1000))

summary(portpirie$SeaLevel)

```
Conclusion















































